Adaptative Computation:\\ An Essay in Memory of John H. Holland

### 2016-02-24 paper.tex
steph's corrections on top of mel's

The emphasis on exploratory models to build intuitions was an
important theme of Holland's work, and he often quoted Eddington's
remark on the occasion of the first experimental test of Einstein's
theory of relativity: ``The contemplation in natural science of a wider
domain than the actual leads to a far better understanding of the
actual'' \cite{Eddington1927}.
%Holland's view of modeling was by no means
%typical.  As just one example of the prevailing view that models are
%for predicting: 
%For example, a textbook on computational modeling offers
% the following definition: 
%``Modeling is the application
%of methods to analyze complex, real-world problems in order to make
%predictions about what might happen with various actions''
%\cite{Shiflet2014}.   This perspective rules out the
%kind of exploratory modeling that interested Holland the most.

@@

ECHO \cite{Holland1999,Hraber1997} was an even more ambitious exploratory model
that Holland and collaborators developed during the 1990s. 
% ECHO's
% scope included the evolution of interacting populations of individual
% agents, which through competition and learning discover symbiotic
% triads such as the famous ant-fly-caterpillar interaction, the Wiksell
% triangle of trading relationships in economics, and the immune
% system's learned ability to distinguish `self' from `other.'
ECHO formalized Holland's idealization of complex adaptive systems
into a runnable computational system where agents evolved external
markers (called tags) and internal preferences, and then used them to
acquire resources and form higher-level aggregate structures, such as
trading relationships, symbiotic groups, trophic cascades, and
interdependent organizations.  ECHO agents sometimes discovered
mimicry and used it to deceive competitors.  In many runs, the model
developed increasingly complex structures and behaviors. % up to a point.
When these runs stabilized, the resulting population of agents was
found to reproduce several well-known patterns observed in nature,
most famously the rank-frequency distribution of species diversity
known as the Preston curve in ecology.  In layman's terms, ECHO
evolved populations where ``most species are rare.''

>>>>

The emphasis on exploratory models to build intuitions was an
important theme of Holland's work, and he often quoted Eddington's
remark on the occasion of the first experimental test of Einstein's
theory of relativity: ``The contemplation in natural science of a wider
domain than the actual leads to a far better understanding of the
actual'' \cite{Eddington1927}.
This view of modeling is not typical.  For many, models
are useful only to the extent that they generate accurate predictions, a
perspective that rules out the very kind of exploratory modeling that
interested Holland the most.
%Holland's view of modeling was by no means
%typical.  As just one example of the prevailing view that models are
%for predicting: 
%For example, a textbook on computational modeling offers
% the following definition: 
%``Modeling is the application
%of methods to analyze complex, real-world problems in order to make
%predictions about what might happen with various actions''
%\cite{Shiflet2014}.   This perspective rules out the
%kind of exploratory modeling that interested Holland the most.

@@

ECHO \cite{Holland1999,Hraber1997} was an even more ambitious exploratory model
that Holland and collaborators developed during the 1990s. 
% ECHO's
% scope included the evolution of interacting populations of individual
% agents, which through competition and learning discover symbiotic
% triads such as the famous ant-fly-caterpillar interaction, the Wiksell
% triangle of trading relationships in economics, and the immune
% system's learned ability to distinguish `self' from `other.'
ECHO formalized Holland's idealization of complex adaptive systems
into a runnable computational system where agents evolved external
markers (called tags) and internal preferences, and then used them to
acquire resources and form higher-level aggregate structures, such as
trading relationships, symbiotic groups, trophic cascades, and
interdependent organizations.  ECHO agents sometimes discovered
mimicry and used it to deceive competitors.  These complex interacting
patterns arose throughout the model's execution, which started with a single
minimal agent.
%In many runs, the model
%developed increasingly complex structures and behaviors. % up to a point.
When the runs stabilized, the resulting population of agents was
found to reproduce several well-known patterns observed in nature,
most famously the rank-frequency distribution of species diversity
known as the Preston curve in ecology.  In layman's terms, ECHO
evolved populations where ``most species are rare.''
